On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-19 TTau_Halpha/kurosawa.tex
th
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@622 25464f21-6099-402c-b30a-640f1d40d38b

\caption{Summary of the standard classical T~Tauri star model parameters.}
\label{tab:std_parameters}

@@

xxxx finished here xxxxx

@@

\subsection{Magnetosphere}

@@

Using the standard parameters (Table~\ref{tab:std_parameters}) for the
central star and the magnetosphere, we examine the dependency of
H$\alpha$ on the temperature ($T_{\mathrm{max}}$) of accretion flow
and the mass accretion rate ($\dot{M}_{\mathrm{acc}}$), as similarly
done by \citet{muzerolle:2001} for H$\beta$. The hot ring temperature
is computed from the available kinetic energy of the free-falling gas
as describe by \citet{hartmann:1994} while \citet{muzerolle:2001} used
the constant hot ring temperature ($8000\,\mathrm{K}$) for most of
their models. The accretion luminosity ($L_{\mathrm{acc}}$) of models
with $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ is
about a half of the total luminosity (without the hot ring) of the
star, and $L_{\mathrm{acc}}$ is proportional to
$\dot{M}_{\mathrm{acc}}$.  The results are placed in
Figure~\ref{fig:atlas_acc_models}. Overall dependency on
$T_{\mathrm{max}}$ and $\dot{M}_{\mathrm{acc}}$ are similar to that of
\citet{muzerolle:2001}. In general, the line strength becomes weaker
as the accretion rate and the temperature become smaller.  The
red-shifted absorption becomes less visible for higher accretion rate
and temperature models in which the flux in the damping wings become
important. Figure~\ref{fig:broadening} shows an example of the effect
of the broadening of H alpha (with $T_{\mathrm{max}}=7500\,\mathrm{K}$
and $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$) due
to the damping constants as described in
section~\ref{sec:radiative-transfer-model}.  Although the maximum flux
of the model with the broadening is almost identical to that of the
model with no damping constant ($\Gamma=0$), a significant increase of
the line flux in both red and blue wings of seen in the model with
broadening. The weak red-shifted absorption component (which is a
signature of the magnetospheric accretion) is weakened or eliminated
by the flux in broadened wing.

Table~\ref{tab:ha_ew_acc_models} shows the EW for the models shown in
Figure~\ref{fig:atlas_acc_models}. About a half of the models shown
the figure agree with the observed EW values of H$\alpha$ from 30 TTS
presented by \citet{alencar:2000} who found the EW of H$\alpha$ ranges
from $\sim 3$~\AA\, to $\sim 160$~\AA, and the mean to be $\sim
55$~\AA. For the lowest $\dot{M}_{\mathrm{acc}}$ models, the EW values
are smaller than the minimum EW observed by \citet{alencar:2000}, and
for some models the EWs are negative. Since the target selection
criteria of \citet{alencar:2000} is not stated in their paper, we can
not conclude that these low $\dot{M}_{\mathrm{acc}}$ models disagree
with the observation. \citet{reipurth:1996}, who measured the EW of 43
TTS and found similar distribution of EW values, mentioned that their
EW measurements are under-representing real samples since the
selection of the targets are based on the strong emission in H$\alpha$
surveys.

The dependency of the profile on the inclination angle ($i$) is
demonstrated in Figure~\ref{fig:acc_inclination_effect}. The model
uses the same parameters as in Figure~\ref{fig:broadening} (with
broadening, $T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$).  The
figure shows that the peak (normalised) flux decreases as the
inclination angle increases. Similarly, the equivalent width also
decreases as the inclination increases. Because of the geometry of the
magnetospheric accretion (c.f. Figure~\ref{fig:config_dipolar}) and of
the presence of the gas with the highest velocity close to the stellar
surface, the highest red-shifted line-of-sight velocity is visible
only at the high inclination angles. This explains the wider
appearance of the profile with $i=80^{\circ}$compared to the
relatively narrow line appearance of the profile with
$i=10^{\circ}$. Although not shown here, a similar dependency on the
inclination angle is found for the models with different temperatures
($T_{\mathrm{max}}=6500$, $8500$, $9500\,\mathrm{K}$).

As seen in the models of \citet{hartmann:1994} and
\citet{muzerolle:2001}, our models also show the blue-shifted peak and
the blue-ward asymmetry caused by the occultation of the accretion
flow by the equatorial disc and the stellar disc. However,
\citealt{alencar:2000} (see their Fig.~9) found a substantial fraction
of the observed H$\alpha$ profiles also shows the {}``red-shifted''
peak, and the P~Cygni profiles which can not be explained by the
magnetospheric accretion model alone. In addition, a recent study by
\citet{appenzeller:2005} showed that the equivalent width of H$\alpha$
from CTTS increases as the inclination angle increases. Our model with
the magnetospheric accretion flow clearly disagrees with their
finding.

@@

  \caption{H$\alpha$ model profiles for wide ranges of mass accretion
    rate ($\dot{M}_\mathrm{acc}$) and temperature
    ($T_{\mathrm{max}}$). The profiles are computed using only
    magnetospheric accretion flow (i.e. no wind/jet).  All the profiles
    are computed using the parameters of the `standard' model
    (Table~\ref{tab:std_parameters}) and inclination $i=55^{\circ}$. The
    temperature (indicated along the vertical axis) of the model increase
    from top to bottom, and the mass accretion rate (indicated by the
    values in $\mathrm{M_{\odot}\,yr^{-1}}$ at the top) increases from
    left to right. The profiles are similar to those of
    \citet{muzerolle:2001}, available on-line
    (http://cfa--www.harvard.edu/cfa/youngstars/models/magnetospheric\_models.html). }

@@

  \caption{The effect of line broadening for H$\alpha$. The model
    computed with a damping constant ($\Gamma$), described in
    section~\ref{sec:radiative-transfer-model} (solid), is compared with
    the model with no damping, $\Gamma=0$ (dashed).  Both models are
    computed with $T_{\mathrm{max}}=7500~\Kelvin$, $i=55^{\circ}$, and the
    standard parameters given in Table~\ref{tab:std_parameters}. The two
    models have similar peak flux levels (around $V \sim 0~\kmps$), but
    the total flux and the EW of the line increased drastically for the
    model with the damping constant. The broad wings extend to $\sim \pm
    800 \kmps$.  The redshifted absorption feature (very weakly) seen in
    the $\Gamma=0$ model is not seen in the model with the broadening.  }

@@

  \caption{Dependency of H$\alpha$ profiles on inclination ($i$).  The
    profiles are computed with the magnetospheric accretion flow using
    the standard parameters given in Table~\ref{tab:std_parameters} and
    $T_{\mathrm{max}}=7500~\Kelvin$. The solid, dotted, and dashed lines
    are for $i=10^{\circ}$, $55^{\circ}$, and $80^{\circ}$
    respectively. As the inclination becomes larger, the peak flux and the
    equivalent width of the line becomes smaller.  Similar dependency is
    seen in the models with the wide ranges of $T_{\mathrm{max}}$ and
    $\beta$. }

@@

\subsection{Bipolar Wind}

@@

As mentioned earlier, the magnetospheric accretion flow model alone
cannot explain some of the features seen in observations. As an
alternative model, we now examine formation of H$\alpha$ in a simple
bipolar wind as described in section \ref{sub:wind_jet}. The basic
wind model parameters introduced earlier are $\gamma$, $\beta$,
$v_{\infty}$, $b$, $\theta_{\mathrm{wind}}$, $T_{\mathrm{wind}}$ and
$\dot{M}_{\mathrm{wind}}$. To minimise the number of free parameters,
we initially adopt $\gamma=0.05$, $v_{\infty}=210\,\kmps$
\citep{appenzeller:2005} and
$\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$.
Further, the angular extent of the wind (the opening angle of the
wind) is assumed to be $\theta_{\mathrm{wind}}=80^{\circ}$. The degree
of collimation is initially chosen to be $b=4$, i.e. the ratio of the
density in polar direction to the density at the edge of the wind
closest to the accretion disc is about
$\left(\cos0^{\circ}/\cos80^{\circ}\right)^{4}\approx10^{3}$, for
given radii. For simplicity, the wind temperature $T_{\mathrm{wind}}$
is assumed to be isothermal\@. Although the line is potentially
sensitive to the temperature structure of the wind, determination of a
self-consistent wind temperature is beyond the scope of this
paper. Readers are refereed to \citet{hartmann:1982} in which the wind
temperature structure is determined by balancing the radiative cooling
rate (assuming optically thin) with the MHD wave heating rate. With
these parameters kept constant, we examine the characteristics of
H$\alpha$ profiles, as a function of the wind acceleration parameter
$\beta$ and the isothermal wind temperature
$T_{\mathrm{wind}}$. Dependency on the other parameters will be
discussed later in this section.

Figure~\ref{fig:atlas_wind_model} shows the profile computed at a mid
inclination angle $i=55^{\circ}$ for the range of the wind temperature
between $7000\,\mathrm{K}$ and $9000\,\mathrm{K}$, and for the range
of $\beta$ between $0.2$ and $2.0$. The P~Cygni profiles are prominent
in the models with lower values of $T_{\mathrm{wind}}$ and $\beta$
(colder and faster acceleration wind). As the value of $\beta$
increases, the position of the blue-shifted absorption component moves
toward the line centre. which can be seen clearly in the
$T_{\mathrm{wind}}=7000\,\mathrm{K}$ models. While the velocity
position of absorption component for the fastest acceleration
($\beta=0.2$) is almost at the terminal velocity ($v_{\infty}\sim
210\,\kmps$), that for the slowest wind acceleration model
($\beta=4.0$) is located close the line centre ($\sim 50\,\kmps$).  As
one can see in Figure~\ref{fig:wind_vr_rho}, the column density of the
gas moving at $v_{\infty}$ increases as the wind acceleration becomes
larger since the gas moving at $v_{\infty}$ becomes closer to the
photosphere. At the location where the absorption occurs, the gas is
moving faster for the models with the faster wind acceleration models;
hence, the location of the absorption component is bluer as $\beta$
becomes smaller (as the wind acceleration faster). The figure also
shows that the intensity of the emission component increases as
$\beta$ increases for a given $T_{\mathrm{wind}}$. This is mainly
because the density near the photosphere increases as the value of
$\beta$ increases (Figure~\ref{fig:wind_vr_rho}); hence, the
emissivity of increases. Unlike the magnetospheric accretion only
models (Figure~\ref{fig:atlas_acc_models}), the inverse P~Cygni
profiles are not seen here, nor the blue-ward asymmetry. Because of
the absorption component in P~Cygni profile, most of the profiles show
red-ward asymmetry except for the low temperature wind models with
high inclinations (not shown here).

@@

\subsection{Disc-wind model of \citet{knigge:1995}}

@@

As an alternative model for the bipolar wind models (with the
combination of the magnetosphere and the accretion disc) shown in
Figures~\ref{fig:atlas_wind_model} and
\ref{fig:atlas_disc_wind_acc_model}, we have also considered the
disc-wind model of \citet{knigge:1995} as described in
section~\ref{sub:model-disc-wind}.  The former resembles the density
structure of \citealt{krasnopolsky:2003} in large scale ($\sim100$~au)
which described in section~\ref{sub:wind_jet}, and the latter in small
scale ($\sim10$~au) of the same model. The main difference between the
two models are in their density structure.  At a given radius, the
density increases toward the polar direction in the former model, but
the density increases toward the equatorial plane for the models
presented here.

@@

  \caption{The H$\alpha$ profiles computed with the disc-wind only
    models. With the wind mass-loss rate fixed at
    $\dot{M}_{\mathrm{dw}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$, the line
    profiles are computed with different combinations of the wind
    acceleration rate ($\beta$) and the isothermal disc-wind temperature
    ($T_{\mathrm{dw}}$). The wind emission increases as the values of
    $\beta$ and $T_{\mathrm{wind}}$ increase ( as the wind becomes
    accelerates slower and as the wind comes warmer.  As in the
    bipolar wind models (Figure~\ref{fig:atlas_wind_model}), the position
    of the wind absorption moves toward the line centre as $\beta$ becomes
    larger. The morphology of the profiles are similar to that of
    observations. }

@@

  \caption{The disc-wind, magnetosphere hybrid model. The same as
    Figure~\ref{fig:atlas_discwind_only}, but these models (solid) also
    include the magnetosphere accretion flow
    ($T_{\mathrm{\max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$). While
    the emission from the magnetosphere dominates for the models with
    smaller $\beta$ and $T_{\mathrm{dw}}$ (the faster acceleration and the
    colder wind models), the emission from the wind dominates the
    profiles for models with larger $\beta$ and $T_{\mathrm{dw}}$ (the
    slower acceleration and the hotter models). For a comparison, the
    profile computed with only the magnetospheric accretion flow (dotted)
    is overlayed.}

@@

First, the disc-wind models without the magnetosphere are considered.
Figure~\ref{fig:atlas_discwind_only} shows the H$\alpha$ profiles
computed using only the disc-wind component (c.f. Figure~\ref{fig:config_discwind})
for the same isothermal disc-wind temperature ($T_{\mathrm{dw}}$)
and the wind acceleration parameter ($\beta$) ranges used for the
bipolar wind models. The parameters used for the central star are
as in section~\ref{sub:Continuum-Source}, and the other disc-wind
parameters are summarised in Table~\ref{tab:std_discwind_parameters}.
The figure shows that the wind emission increases as $\beta$ and $T_{\mathrm{dw}}$
increase. Similar to the bipolar wind models, the location of the
P-Cyg absorption component moves toward the line centre as $\beta$
increases. The morphology of the profile exhibited by the model changes
from FU-Ori type for the models with small $\beta$ to the stronger
emission types which are more commonly seen in the observed H$\alpha$
(c.f.~\citealt{reipurth:1996}) as the value of $\beta$
increases.  Although not shown here, we find that the profile shapes
are relatively insensitive to the value of the wind collimation parameter
$d$. The dependency on the disc wind mass-loss rate of the profile
shape is similar to that of the collimated wind models
(Figure~\ref{fig:effect_mdot_disc_wind_acc}).

We now consider the models computed with the combination of the
disc-wind and the magnetospheric accretion components. The
parameters used for the magnetosphere are set to be same as in the
case for Figure~\ref{fig:acc_inclination_effect}, and are kept
constant for simplicity. The mass-loss rate of the disc wind is set to
$\dot{M}_{\mathrm{dw}}=10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$
(i.e.~$\mu=\dot{M}_{\mathrm{dw}}/\dot{M}_{\mathrm{acc}}=0.1$).
Figure~\ref{fig:atlas_discwind_acc_model} shows the models profiles
computed using the same ranges of the disc-wind temperature and the
wind acceleration parameter as in the previous cases.

As in the disc wind only models
(Figure~\ref{fig:atlas_discwind_only}), the location of the absorption
component moves toward the line centre as the value of $\beta$
increases (the wind acceleration is slower).  The figure shows that in
the models with smaller $\beta$ and $T_{\mathrm{dw}}$, the line
emission from the magnetosphere dominates. On the other hand, in the
models with larger $\beta$ and $T_{\mathrm{dw}}$, the emission from
the disc wind dominates. Compared to the observation of
\citet{reipurth:1996} and \citet{alencar:2000}, the lines are too
strong for
$\left(T_{\mathrm{dw}},\,\beta\right)=\left(9000\,\mathrm{K},\,2.0\right)$
and $\left(9000\,\mathrm{K},\,1.0\right)$ models.

Although the models in Figure~\ref{fig:atlas_discwind_acc_model} are
computed with limited ranges of  $T_{\mathrm{dw}}$ and $\beta$, the
resulting profiles exhibit the wide varieties of line shapes which are
very similar to the types of H$\alpha$ profiles seen in the
observations (c.f.~\citealt{reipurth:1996}).  The shape of the
primary and secondary peaks in the profiles are slightly {}``rounder''
compared to those of the accretion, collimated wind, disc hybrid model
shown in Figure~\ref{fig:atlas_disc_wind_acc_model}.  The rounder
peaks are caused by the rotation of the wind. The profiles from the
disc-wind models are affected more by the rotational motion of the
wind since the gas is more concentrated toward the equator, but in the
models with the collimated wind the gas is concentrated more in the
polar directions.

We note the ambiguity seen between some of the disc-wind only models
and the disc-wind, magnetosphere hybrid models. For example, the disc-wind
only model with $\left(T_{\mathrm{wind}},\,\beta\right)=\left(8000\,\mathrm{K},\,1.0\right)$
in Figure~\ref{fig:atlas_discwind_only} and the disc-wind model
with $\left(T_{\mathrm{wind}},\,\beta\right)=\left(9000\,\mathrm{K},\,0.5\right)$
in Figure~\ref{fig:atlas_discwind_acc_model} give very similar profiles.
Given such an observed profile, it is difficult to distinguish if
the emission is dominated by the accretion or the wind unless some
other lines, not affected associated with wind or magnetosphere, are
used simultaneously.

General characteristics of the profiles from the disc-wind, magnetosphere
hybrid model (Figure~\ref{fig:atlas_discwind_acc_model}) are very
similar to those of the disc, collimated wind, magnetosphere hybrid
model (Figure~\ref{fig:atlas_disc_wind_acc_model}). The main difference
between these two sets of models is in the ration of the mass-loss rate to the mass
accretion rate, $\mu$. While the models presented in this section
use $\mu=0.1$ which is consistent with the MHD jet models (e.g.~\citealt{koenigl:2000}),
the models in Figure~\ref{fig:atlas_disc_wind_acc_model} use much
smaller value ($\mu=0.01$).   The dependency of the model profiles on
the inclination angle for the two different wind models will be discussed
in the next section.

>>>>

\caption{Summary of the reference classical T~Tauri star model parameters.}
\label{tab:std_parameters}

@@

@@

\subsection{Magnetosphere only models}

@@

Using the reference model parameters (Table~\ref{tab:std_parameters}) for the
central star and the magnetosphere, we examine the dependency of
H$\alpha$ on the temperature ($T_{\mathrm{max}}$) of accretion flow
and the mass accretion rate ($\dot{M}_{\mathrm{acc}}$), as
done by \citet{muzerolle:2001} for H$\beta$. We note that here we have
 computed the hot ring temperature self-consistently, whereas \citet{muzerolle:2001} used
a constant hot ring temperature ($8000\,\mathrm{K}$) for most of
their models.  The accretion luminosity ($L_{\mathrm{acc}}$) for models
with $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$ is
about a half of the photospheric luminosity.

H$\alpha$ line profiles for a range of $T_{\mathrm{max}}$ and
$\dot{M}_{\mathrm{acc}}$ are presented in  Figure~\ref{fig:atlas_acc_models}. 
The  overall dependency on $T_{\mathrm{max}}$ and $\dot{M}_{\mathrm{acc}}$ is similar to that of
\citet{muzerolle:2001}. In general, the line strength weakens
as the accretion rate and the temperature drop.  The
red-shifted absorption becomes less visible for the higher accretion rate
and higher temperature models as it becomes filled-in by the stronger Stark-broadened
line-wing emission.

We demonstrate this effect in Figure~\ref{fig:broadening}, which shows an example 
for an H$\alpha$  model with $T_{\mathrm{max}}=7500\,\mathrm{K}$
and $\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$, both with and without
damping.  Although the maximum flux of the model with broadening is almost identical to that of the
model with no damping constant ($\Gamma=0$), a significant increase of
the line flux in both red and blue wings of seen. The weak red-shifted
absorption component (which is a signature of the magnetospheric accretion)
is weakened or eliminated by the flux in broadened wing.

Table~\ref{tab:ha_ew_acc_models} shows the EW for the models in
Figure~\ref{fig:atlas_acc_models}.  The EWs for half the models
fall within the range of EWs ($\sim 3$~\AA\, to $\sim 160$~\AA)
measured by \citet{alencar:2000}. The equivalent widths of models 
with the lowest mass accretion rates fall below the minimum EW
observed by  \citet{alencar:2000}, while several models would
be designated as weak-lined T~Tauri stars using the traditional
10\,\AA\ cut-off.

xxxx might want to talk about line widths as an accretion 
indicator here as well (white and basri).

The dependency of the line profile on inclination angle ($i$) is
demonstrated in Figure~\ref{fig:acc_inclination_effect}. The model
has  $T_{\mathrm{max}}=7500\,\mathrm{K}$ and
$\dot{M}_{\mathrm{acc}}=10^{-7}\,\mathrm{M_{\sun}\, yr^{-1}}$).  The
figure shows that the peak (normalised) flux decreases as the
inclination angle increases, as does the EW.

xxxx check reasons for inclination dependence xxxx

Because of the geometry of the
magnetospheric accretion (c.f. Figure~\ref{fig:config_dipolar}) and of
the presence of the gas with the highest velocity close to the stellar
surface, the highest red-shifted line-of-sight velocity is visible
only at the high inclination angles. This explains the wider
appearance of the profile with $i=80^{\circ}$compared to the
relatively narrow line appearance of the profile with
$i=10^{\circ}$.

A similar EW dependence on
inclination angle is found for the models with different 
magnetospheric temperatures ($T_{\mathrm{max}}=6500$,
$8500$, $9500\,\mathrm{K}$).

Our models show a blue-shifted line peak and
a blueward asymmetry caused by occultation of the accretion
flow by the stellar photosphere (see also \citet{hartmann:1994} and
\citet{muzerolle:2001}). However, \citealt{alencar:2000} (see their Fig.~9)
found a substantial fraction
of the observed H$\alpha$ profiles show a red-shifted
peak. Furthermore, a recent study by
\citet{appenzeller:2005} demonstrated  that the equivalent width of H$\alpha$
from CTTS {\em increases} as the inclination angle increases. 
Clearly a magnetospheric accretion only model cannot explain the
full properties of the H$\alpha$ profiles in CTTS.

@@

  \caption{H$\alpha$ model profiles for wide ranges of mass accretion
    rate ($\dot{M}_\mathrm{acc}$) and temperature
    ($T_{\mathrm{max}}$). The profiles are computed using only the
    magnetospheric accretion flow (i.e. no outflow).  All the profiles
    are computed using the parameters of the reference model
    (Table~\ref{tab:std_parameters}) and inclination $i=55^{\circ}$. The
    temperature (indicated along the vertical axis) of the model increases
    from top to bottom, and the mass accretion rate (indicated by the
    values in $\mathrm{M_{\odot}\,yr^{-1}}$ along the top) increases from
    left to right. }

@@

  \caption{The effect of line broadening on H$\alpha$. A model
    computed with a damping constant ($\Gamma$), described in
    section~\ref{sec:radiative-transfer-model} (solid), is compared with
    a model with no damping, $\Gamma=0$ (dashed).  Both models are
    computed with $T_{\mathrm{max}}=7500~\Kelvin$, $i=55^{\circ}$, and the
    reference parameters given in Table~\ref{tab:std_parameters}. The two
    models have similar peak flux levels (around $V \sim 0~\kmps$), but
    the total flux and the EW of the line increased drastically for the
    model with the damping constant. The broad wings extend to $\sim \pm
    800 \kmps$.  The redshifted absorption feature (very weakly) seen in
    the $\Gamma=0$ model is not seen in the model with the broadening.  }

@@

  \caption{The dependency of the H$\alpha$ profile on inclination ($i$).  The
    profiles are computed with a magnetospheric accretion flow only using
    the reference parameters given in Table~\ref{tab:std_parameters} and
    $T_{\mathrm{max}}=7500~\Kelvin$. The solid, dotted, and dashed lines
    are for $i=10^{\circ}$, $55^{\circ}$, and $80^{\circ}$
    respectively. As the inclination becomes larger, the peak flux and the
    equivalent width of the line becomes smaller.  }

@@

\subsection{Bipolar wind only models}

@@

As mentioned above, the MA model alone
cannot explain the observations. As an
alternative model, we now examine formation of H$\alpha$ in a simple
bipolar wind as described in section \ref{sub:wind_jet}. The basic
wind model parameters introduced earlier are $\gamma$, $\beta$,
$v_{\infty}$, $b$, $\theta_{\mathrm{wind}}$, $T_{\mathrm{wind}}$ and
$\dot{M}_{\mathrm{wind}}$. We fix several of these parameters
at plausible values:  $\gamma=0.05$, $v_{\infty}=210\,\kmps$
\citep{appenzeller:2005} and
$\dot{M}_{\mathrm{wind}}=10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$.
Further, the angular extent of the wind (the opening angle of the
wind) is assumed to be $\theta_{\mathrm{wind}}=80^{\circ}$. The degree
of collimation is initially chosen to be $b=4$, i.e. the ratio of the
density in polar direction to the density at the edge of the wind
closest to the accretion disc is about
$\left(\cos0^{\circ}/\cos80^{\circ}\right)^{4}\approx10^{3}$
Although the line is potentially
sensitive to the temperature structure of the wind, determination of a
self-consistent wind temperature is beyond the scope of this
paper. Readers are refereed to \citet{hartmann:1982} in which the wind
temperature structure is determined by balancing the radiative cooling
rate (assuming optically thin) with the MHD wave heating rate.
We pragmatically assume that the wind is isothermal at $T_{\mathrm{wind}}$.

Initially we examine the characteristics of the
H$\alpha$ profile  as a function of the wind acceleration parameter
$\beta$ and the isothermal wind temperature
$T_{\mathrm{wind}}$.  Figure~\ref{fig:atlas_wind_model} shows profiles computed at  $i=55^{\circ}$
 for a range of wind temperatures
between $7000\,\mathrm{K}$ and $9000\,\mathrm{K}$, and  $\beta$ between $0.2$ and $2.0$.

Classical P~Cygni profiles are prominent
in the models with lower values of $T_{\mathrm{wind}}$ and $\beta$
(a colder and faster accelerating wind). As the value of $\beta$
increases, the position of the blue-shifted absorption component moves
toward the line centre -- this can be seen clearly in the
$T_{\mathrm{wind}}=7000\,\mathrm{K}$ models. The velocity position
 of the absorption component of the fastest accelerating wind
($\beta=0.2$) is almost at the terminal velocity ($v_{\infty}\sim
210\,\kmps$), that for the slowest wind acceleration model
($\beta=4.0$) is located close the line centre ($\sim 50\,\kmps$).

 As
one can see in Figure~\ref{fig:wind_vr_rho}, the column density of the
gas moving at $v_{\infty}$ increases as the wind acceleration becomes
larger since the gas moving at $v_{\infty}$ becomes closer to the
photosphere. At the location where the absorption occurs, the gas is
moving faster for the models with the faster wind acceleration models;
hence, the location of the absorption component is bluer as $\beta$
becomes smaller (as the wind acceleration faster). The figure also
shows that the intensity of the emission component increases as
$\beta$ increases for a given $T_{\mathrm{wind}}$. This is mainly
because the density near the photosphere increases as the value of
$\beta$ increases (Figure~\ref{fig:wind_vr_rho}); hence, the
emissivity of increases. Unlike the magnetospheric accretion only
models (Figure~\ref{fig:atlas_acc_models}), the inverse P~Cygni
profiles are not seen here, nor the blue-ward asymmetry. Because of
the absorption component in P~Cygni profile, most of the profiles show
red-ward asymmetry except for the low temperature wind models with
high inclinations (not shown here).

@@

\subsection{The disc-wind model}

@@

In this section we examine profiles produced using our alternative
kinematic model of the outflow: the disc wind. This type of outflow
has been predicted by several MHD calculations (e.g. \citet{koenigl:2000};
 \citealt{krasnopolsky:2003}), and broadly consists of a wind with
a latitudinal density-dependence with density increasing towards
the equator (see~\ref{sub:model-disc-wind}), in contrast with the 
bipolar stellar wind.

@@

  \caption{H$\alpha$ profiles computed with the disc-wind only.
     The wind mass-loss rate is fixed at
    $\dot{M}_{\mathrm{dw}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$, and the line
    profiles are computed with different combinations of the wind
    acceleration rate ($\beta$) and isothermal disc-wind temperature
    ($T_{\mathrm{dw}}$). The wind emission grows as the values of
    $\beta$ and $T_{\mathrm{wind}}$ increase.   As in the
    bipolar wind models (Figure~\ref{fig:atlas_wind_model}), the position
    of the wind absorption moves toward the line centre as $\beta$ becomes
    larger.     }

@@

  \caption{The disc-wind, magnetosphere hybrid model. These models
  (solid lines) have the  same parameters as
    Figure~\ref{fig:atlas_discwind_only}, but  also
    include the magnetospheric accretion flow
    ($T_{\mathrm{\max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-8}\,\mathrm{M_{\sun}\,yr^{-1}}$). While
    the emission from the magnetosphere dominates for the models with
    smaller $\beta$ and $T_{\mathrm{dw}}$, the emission from the wind dominates the
    profiles for models with larger $\beta$ and $T_{\mathrm{dw}}$.  The contribution
    of the magnetosphere can be judged from the profiles computed with
    the magnetosphere only (dotted lines).}

@@

Firstly, we consider the disc-wind models without the magnetosphere.
Figure~\ref{fig:atlas_discwind_only} shows the H$\alpha$ profiles
computed using only the disc-wind component (c.f. Figure~\ref{fig:config_discwind})
for the same isothermal disc-wind temperature ($T_{\mathrm{dw}}$)
and the wind acceleration parameter ($\beta$) ranges used for the
bipolar wind models (c.f. xxxx cite bipolar figure here). The parameters used for the central star are
as in section~\ref{sub:Continuum-Source}, and the disc-wind
parameters are summarised in Table~\ref{tab:std_discwind_parameters}.
The figure shows that the wind emission increases as $\beta$ and $T_{\mathrm{dw}}$
increase. The $\beta$ dependence arises simply because for large $\beta$
there is a greater mass of wind material close to the star.
As with  the bipolar wind models, the location of the
P-Cyg absorption component moves toward the line centre as $\beta$
increases.

The morphology of the profile exhibited by the model changes
from FU-Ori type for the models with small $\beta$ to the stronger
emission types which are more commonly seen in the observed H$\alpha$
(c.f.~\citealt{reipurth:1996}) as the value of $\beta$
increases.  We find that the profile shapes
are relatively insensitive to the value of the wind collimation parameter
$d$. The dependency on the disc wind mass-loss rate of the profile
shape is similar to that of the bipolar wind models
(Figure~\ref{fig:effect_mdot_disc_wind_acc}).

We now consider the models computed with a combination of the
disc-wind and the magnetospheric accretion components. The
parameters used for the magnetosphere are set to be same as in the
case for Figure~\ref{fig:acc_inclination_effect}.  
The mass-loss rate of the disc wind is set to
$\dot{M}_{\mathrm{dw}}=10^{-8}\,\mathrm{M_{\sun}\, yr^{-1}}$
(i.e.~$\mu=\dot{M}_{\mathrm{dw}}/\dot{M}_{\mathrm{acc}}=0.1$).
Figure~\ref{fig:atlas_discwind_acc_model} shows the models profiles
computed using the same ranges of the disc-wind temperature and the
wind acceleration parameter as in the previous cases.

As in the disc-wind only models
(Figure~\ref{fig:atlas_discwind_only}), the location of the absorption
component moves toward the line centre as the value of $\beta$
increases.  The figure shows that in
the models with smaller $\beta$ and $T_{\mathrm{dw}}$, the line
emission from the magnetosphere dominates, while this situation reverses for the
models with larger $\beta$ and $T_{\mathrm{dw}}$. The model with
$\left(T_{\mathrm{dw}},\,\beta\right)=\left(9000\,\mathrm{K},\,2.0\right)$
results in an H$\alpha$ profile that is far too strong to be compatible
 with observations (\citet{reipurth:1996} and \citet{alencar:2000}).

Although the models in Figure~\ref{fig:atlas_discwind_acc_model} are
computed with limited ranges of  $T_{\mathrm{dw}}$ and $\beta$, the
resulting profiles exhibit a wide variety of line profile shapes,
many of  which are similar to the types of H$\alpha$ profiles seen in the
observations (c.f.~\citealt{reipurth:1996}).   The shape of the
primary and secondary peaks in the profiles are slightly {}``rounder''
in comparison to those of the accretion, bipolar wind, disc hybrid model
shown in Figure~\ref{fig:atlas_disc_wind_acc_model}.  The rounder
peaks are the result of wind rotation.

xxxx think of a way of talking about rotation in comparison between
the disc wind model and the bipolar wind model xxxx

An examination of the profiles in Figures~\ref{fig:atlas_discwind_only} and
Figure~\ref{fig:atlas_discwind_acc_model} reveals some of problems
with uniqueness that must inevitably occur when trying to assign
an individual profile as wind or accretion dominated i.e. very similar profiles may arise
from very different circumstellar geometries. For example, the disc-wind
only model with $\left(T_{\mathrm{wind}},\,\beta\right)=\left(8000\,\mathrm{K},\,1.0\right)$
(Figures~\ref{fig:atlas_discwind_only}) and the  accretion dominated model with
 $\left(T_{\mathrm{wind}},\,\beta\right)=\left(9000\,\mathrm{K},\,0.5\right)$
in (Figure~\ref{fig:atlas_discwind_acc_model}) give very similar profiles.

xxxx paragraph below may need changing after new models are added xxxx

The general characteristics of the profiles from the disc-wind, magnetosphere
hybrid model (Figure~\ref{fig:atlas_discwind_acc_model}) are very
similar to those of the disc, bipolar wind, magnetosphere hybrid
model (Figure~\ref{fig:atlas_disc_wind_acc_model}). The main difference
between these two sets of models is in the ratio of the mass-loss rate to the mass
accretion rate, $\mu$. While the models presented in this section
use $\mu=0.1$ which is consistent with the MHD jet models (e.g.~\citealt{koenigl:2000}),
the models in Figure~\ref{fig:atlas_disc_wind_acc_model} use much
smaller value ($\mu=0.01$).   The dependency of the model profiles on
the inclination angle for the two different wind models will be discussed
in the next section.
